LME-MPM applied to quasi-brittle fracture.

### 2020-06-30 2020_EFM_MPM_Eigensoftening.tex
Version final

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of presenting a suitable alternative
  to the existing fracture algorithms developed for \PNA{such a methodology}.
  %\acrshort{mpm}. 
  The good fitting in the predictions made by the
  eigensoftening algorithm against both analytical and experimental
  results proofs the \PNA{excellent} performance of the method when challenging
  applications are to modeled.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
\DM{The failure mechanisms of quasi-brittle materials and the  crack propagation have been studied from different fields such as materials science, structural engineering, earth science and petroleum engineering . The prediction of crack propagation has been approached with different numerical techniques. The \acrfull{fem} is widely used in industrial application. However,} the presence of cracks is a violation of the continuity requirement of the \acrshort{fem}. To overcome this issue, numerous numerical
artifacts have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, or \PNA{to the gradual insertion of}
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries. To alleviate these shortcomings of the
\acrshort{fem}, the \acrfull{xfem} in conjunction with level set
technique was introduced by Belytschko {\it et al.}
\cite{Belytschko_03} to handle arbitrary crack paths level
set. \MMP{Some relevant contributions to the \acrshort{xfem} are the
modeling micro-scale crack propagation \cite{HUANG_2018}, the
introduction of a cohesive zone model \cite{NAGASHIMA_2016}, and more
recently improving it capability to predict the crack path in 3D cases
\cite{YUN_2019}.}

The simulation of fracture propagation in a more accurate and effective
way can be considered as one of the \PNA{ongoing goals} of the development of novel spatial discretisation methods like meshfree techniques. Some
examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
\acrfull{efgm} \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013}, the
\acrfull{sph} \cite{Wang_2020,Wang_2019}, the \acrfull{otm} \cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or Peridynamics
\cite{HA_2011,RABCZUK_2017} among others. \\

\MMP{Concerning} \acrshort{mpm}, fracture \PNA{has been treated numerically through two different ways}. On the one hand, the ``CRAcks with Material Points (CRAMP)'' is proposed~\cite{Nairn_2003,Nairn_2006}. \PNA{The basis of this methodology consist of the removal of} the restriction
of the single-valued velocity field close to the crack through, at least, two sets of nodes. In this method, different labels are assigned to the material points
and nodes to distinguish if they are in the same side of the crack or
not. Under this approach, crack surface is described with line segments
in 2D and triangle patches in 3D cases. The chosen criteria for crack
propagation is based on such parameters as energy release rate
analysed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. 
On the other hand, the \PNA{foundation} of the second approach is the introduction of failed
material points to describe the crack evolution. In these methods, the
formation of failed points describes the nucleation of cracks, and
thereafter its propagation and branching. Consequently, the position
of the crack does not need to be explicitly stated. This fact represents
significant advantages over the ``CRAMP''. Under this approach, the
prediction of failure evolution is computed with a decohesion model,
which has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} and
Schreyer {\it et al.}\cite{Schreyer_2002}. Successful simulations are provided within the framework of the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}.\\

Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
crack sets are approximated by means of eigen-deformations, which
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle fracture
was carried out by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. \PNA{In this technique, the concept of ``erosion'' of
the material point is depicted, being the material point able to remain intact or become completely failed or eroded, when no loading
capacity is bore}. This method has been successfully applied to simulate high
complex phenomena such as dynamic fragmentation of metals
\cite{Li_2015}. Recently, Zhang {\it et al.}
\cite{Zhang_EE_2020} adopted the eigenerosion to resolve the dynamic
fracture of brittle materials in the \acrshort{mpm}
framework. Nonetheless, when quasi-brittle materials are simulated
though this approach, the results exhibit an overestimation of tensile
stress and strain peaks. Furthermore, since the work of Zhang {\it et al.}
\cite{Zhang_EE_2020} is made with conventional \acrshort{mpm}/\acrshort{gimp}, stress oscillations are to be faced.\\

To overcome the limitations observed in the EigenMPM
\cite{Zhang_EE_2020}, the present research proposes the eigensoftening
algorithm developed by Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} for the \acrshort{otm} framework and
engineered for quasi-brittle materials. Inspired in the concept of the
crack band model \cite{Bazant83}, since energy dissipation is performed through the softened (or failed) volume, this methodology is able to capture the gradual rather than abrupt
dissipation of the fracture energy. \PNA{Moreover, in order to mitigate stress
oscillations as well as to deal with tensile instabilities, such those that
occurs when particles crosses an element boundary, the work of Molinos~\etal~\cite{Molinos2020} is followed. The traditional \acrshort{mpm} technique is enhanced through the \acrfull{lme} approximation technique
\cite{Arroyo2006} for the spatial discretisation as well as a explicitly \acrshort{mpm} \acrfull{npc} scheme is derived for the time discretisation.}\\

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials.\DM{The
  evaluation of both approaches \cite{Navas_2017_ES}}\MMP{against
  experimental measures}\DM{show that the eigenerosion algorithm significantly overestimates the tensile stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately}. Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

The key idea behind the eigenerosion algorithm is the computation of the
energy-release rate attendant to the failure of material point $p$,
\begin{align}
  \label{eq:energy-release-EE}
&G_p^{k+1} = \frac{C_{\epsilon}}{m_p^{k+1}}  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q W_q^{k+1}\\
  \label{eq:mass-EE}
&m_p^{k+1} =  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q  
\end{align}
where $B_{\epsilon}(x_p^{k+1})$ is a $^n$dimensional sphere of radius $\epsilon$
centered at $x_p^{k+1}$. The particles which lie under this sphere are
known as the $\epsilon$-neighborhood of the material point (see
\cite{Pandolfi_2012}). This concept is conveniently sketched in
Figure \ref{fig:Failed-particles}.
%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure-Particle-failed}
  \caption{Scheme of a linear cohesive law, where the shaded area is
    $G_f$, $f_t$ is the tensile strength, and $w_c$ is the critical
    opening displacement.}
  \label{fig:Failed-particles}
\end{figure}
%%%%%%%%%%%
Other parameters are the mass of the neighborhood $m_p^{k+1}$, the
current free-energy density per unit mass  $W_q^{k+1}$ and the
normalizing constant $C_{\epsilon}$, \PNA{which also defines the
  $\epsilon$-neighborhood size, being $\epsilon = C_\epsilon h$}, \MMP{ and
  $h$ is a measure of the distance between nodes.}

@@

initialization was adopted. \MMP{Particularly the maximum principal stress

@@

$\epsilon$-neighborhood of the material point $\vec{x}_p^{k+1}$ as}

@@

$\epsilon$-neighborhood,\DM{defined similarly to the eigenerosion

@@

  leads to a definition of the   averaged strain energy} in terms of the averaged

@@

variable and the volume for each material point $p$. \PNA{$\chi_p$ is any function, taking

@@

material points), that relates the crack opening and the residual strength. To calculate it, typical shapes of cohesive laws such as linear, bilinear or exponential, may be employed.} For the case of a linear softening such the sketched one in the Figure \ref{fig:Damage-ft-wc}, $\chi_p$ is computed as,

@@

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
\PNA{The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology.} Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, \PNA{the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error} prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead of with the belonging element as heretofore has been
done.
\MMP{Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest nodes are defined, the particle
possesses automatically a list of \textit{tributary nodes}. This has two
immediate benefits: the first one is that shape functions with global
support as \acrshort{lme} are easier to be implemented,
and the second one is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, it is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\DM{The proposed approach to overcome the two main shortcomings of
  EighenMPM \cite{Zhang_EE_2020} has been evaluated with ``three
  benchmark tests''}.
The first one deals with the presence of stress
instabilities provoked by the spatial discretisation, appearing even when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}.\MMP{A proof
of it is exposed in Sections \ref{sec:3.2} and \ref{sec:3.3}, where experimental results
are compared \PNA{with numerical ones, being the eigensoftening technique the one employed in this research for the
computations.} In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.}

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation. The application
consists of a square plate of size\MMP{$H = 1\ [L] $} containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \MMP{\gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current simulation will not be tested against experimental results, where
scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}.} In
\acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025 $[L]$. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

>>>>

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of presenting a suitable alternative
  to the existing fracture algorithms developed for such a methodology.
  %\acrshort{mpm}. 
  The good fitting in the predictions made by the
  eigensoftening algorithm against both analytical and experimental
  results proofs the excellent performance of the method when challenging
  applications are to modeled.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
The failure mechanisms of quasi-brittle materials and the  crack propagation have been studied from different fields such as materials science, structural engineering, earth science and petroleum engineering . The prediction of crack propagation has been approached with different numerical techniques. The \acrfull{fem} is widely used in industrial application. However, the presence of cracks is a violation of the continuity requirement of the \acrshort{fem}. To overcome this issue, numerous numerical
artifacts have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, or to the gradual insertion of
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries. To alleviate these shortcomings of the
\acrshort{fem}, the \acrfull{xfem} in conjunction with level set
technique was introduced by Belytschko {\it et al.}
\cite{Belytschko_03} to handle arbitrary crack paths level
set. Some relevant contributions to the \acrshort{xfem} are the
modeling micro-scale crack propagation \cite{HUANG_2018}, the
introduction of a cohesive zone model \cite{NAGASHIMA_2016}, and more
recently improving it capability to predict the crack path in 3D cases
\cite{YUN_2019}.

The simulation of fracture propagation in a more accurate and effective
way can be considered as one of the ongoing goals of the development of novel spatial discretisation methods like meshfree techniques. Some
examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
\acrfull{efgm} \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013}, the
\acrfull{sph} \cite{Wang_2020,Wang_2019}, the \acrfull{otm} \cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or Peridynamics
\cite{HA_2011,RABCZUK_2017} among others. \\

Concerning \acrshort{mpm}, fracture has been treated numerically through two different ways. On the one hand, the ``CRAcks with Material Points (CRAMP)'' is proposed~\cite{Nairn_2003,Nairn_2006}. The basis of this methodology consist of the removal of the restriction
of the single-valued velocity field close to the crack through, at least, two sets of nodes. In this method, different labels are assigned to the material points
and nodes to distinguish if they are in the same side of the crack or
not. Under this approach, crack surface is described with line segments
in 2D and triangle patches in 3D cases. The chosen criteria for crack
propagation is based on such parameters as energy release rate
analysed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. 
On the other hand, the foundation of the second approach is the introduction of failed
material points to describe the crack evolution. In these methods, the
formation of failed points describes the nucleation of cracks, and
thereafter its propagation and branching. Consequently, the position
of the crack does not need to be explicitly stated. This fact represents
significant advantages over the ``CRAMP''. Under this approach, the
prediction of failure evolution is computed with a decohesion model,
which has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} and
Schreyer {\it et al.}\cite{Schreyer_2002}. Successful simulations are provided within the framework of the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}.\\

Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
crack sets are approximated by means of eigen-deformations, which
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle fracture
was carried out by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. In this technique, the concept of ``erosion'' of
the material point is depicted, being the material point able to remain intact or become completely failed or eroded, when no loading
capacity is bore. This method has been successfully applied to simulate high
complex phenomena such as dynamic fragmentation of metals
\cite{Li_2015}. Recently, Zhang {\it et al.}
\cite{Zhang_EE_2020} adopted the eigenerosion to resolve the dynamic
fracture of brittle materials in the \acrshort{mpm}
framework. Nonetheless, when quasi-brittle materials are simulated
though this approach, the results exhibit an overestimation of tensile
stress and strain peaks. Furthermore, since the work of Zhang {\it et al.}
\cite{Zhang_EE_2020} is made with conventional \acrshort{mpm}/\acrshort{gimp}, stress oscillations are to be faced.\\

To overcome the limitations observed in the EigenMPM
\cite{Zhang_EE_2020}, the present research proposes the eigensoftening
algorithm developed by Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} for the \acrshort{otm} framework and
engineered for quasi-brittle materials. Inspired in the concept of the
crack band model \cite{Bazant83}, since energy dissipation is performed through the softened (or failed) volume, this methodology is able to capture the gradual rather than abrupt
dissipation of the fracture energy. Moreover, in order to mitigate stress
oscillations as well as to deal with tensile instabilities, such those that
occurs when particles crosses an element boundary, the work of Molinos~\etal~\cite{Molinos2020} is followed. The traditional \acrshort{mpm} technique is enhanced through the \acrfull{lme} approximation technique
\cite{Arroyo2006} for the spatial discretisation as well as a explicitly \acrshort{mpm} \acrfull{npc} scheme is derived for the time discretisation.\\

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. The
  evaluation of both approaches \cite{Navas_2017_ES} against
  experimental measures show that the eigenerosion algorithm significantly overestimates the tensile stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

The key idea behind the eigenerosion algorithm is the computation of the
energy-release rate attendant to the failure of material point $p$,
\begin{align}
  \label{eq:energy-release-EE}
&G_p^{k+1} = \frac{C_{\epsilon}}{m_p^{k+1}}  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q W_q^{k+1}\\
  \label{eq:mass-EE}
&m_p^{k+1} =  \sum_{x_q^{k+1} \in
  B_{\epsilon}(x_p^{k+1})} m_q  
\end{align}
where $B_{\epsilon}(x_p^{k+1})$ is a $^n$dimensional sphere of radius $\epsilon$
centered at $x_p^{k+1}$. The particles which lie under this sphere are
known as the $\epsilon$-neighborhood of the material point (see
\cite{Pandolfi_2012}). This concept is conveniently sketched in
Figure \ref{fig:Failed-particles}.
%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Figure-Particle-failed}
  \caption{Scheme of a linear cohesive law, where the shaded area is
    $G_f$, $f_t$ is the tensile strength, and $w_c$ is the critical
    opening displacement.}
  \label{fig:Failed-particles}
\end{figure}
%%%%%%%%%%%
Other parameters are the mass of the neighborhood $m_p^{k+1}$, the
current free-energy density per unit mass  $W_q^{k+1}$ and the
normalizing constant $C_{\epsilon}$, which also defines the
  $\epsilon$-neighborhood size, being $\epsilon = C_\epsilon h$, and
  $h$ is a measure of the distance between nodes.

@@

initialization was adopted. Particularly the maximum principal stress

@@

$\epsilon$-neighborhood of the material point $\vec{x}_p^{k+1}$ as

@@

$\epsilon$-neighborhood, defined similarly to the eigenerosion

@@

  leads to a definition of the   averaged strain energy in terms of the averaged

@@

variable and the volume for each material point $p$. $\chi_p$ is any function, taking

@@

material points), that relates the crack opening and the residual strength. To calculate it, typical shapes of cohesive laws such as linear, bilinear or exponential, may be employed. For the case of a linear softening such the sketched one in the Figure \ref{fig:Damage-ft-wc}, $\chi_p$ is computed as,

@@

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology. Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead of with the belonging element as heretofore has been
done.
Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest nodes are defined, the particle
possesses automatically a list of \textit{tributary nodes}. This has two
immediate benefits: the first one is that shape functions with global
support as \acrshort{lme} are easier to be implemented,
and the second one is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, it is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The proposed approach to overcome the two main shortcomings of
EighenMPM \cite{Zhang_EE_2020} has been evaluated with ``three
benchmark tests''.
The first one deals with the presence of stress
instabilities provoked by the spatial discretisation, appearing even when the \acrshort{gimp} shape functions are employed. To overcome it, \acrshort{lme}
approximants are introduced as an alternative to the existing interpolation techniques. Its
performance in mitigating spurious stress oscillations under the \acrshort{mpm}~\cite{Wobbes2020,Molinos2020} framework helps to enhance
notoriously the quality of the results as it is observed in Section
\ref{sec:3.1}, where both interpolation techniques
are compared by carrying out an eigenerosion simulation. The second
limitation is concerned to its inability to simulate properly quasi-brittle
fracture \cite{Navas_2018_ES}. It can be solved through the
eigensoftening algorithm described in Section \ref{sec:2.3}. A proof
of it is exposed in Sections \ref{sec:3.2} and \ref{sec:3.3}, where experimental results
are compared with numerical ones, being the eigensoftening technique the one employed in this research for the
computations. In the first case comparing with a semicircular bending
test and in the second case with a drop-weight impact test.

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented hereinafter is to assess the
capability of the \acrshort{lme} approximants to improve the result
\textit{versus} the standard linear interpolation and \acrshort{ugimp}
\cite{Bardenhagen2004}. The application
consists of a square plate of size $H = 1\ [L] $ containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06
$[M] [L]^{-1} [T]^{-2}$, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001 $[L]^{2} [T]^{-2}$. The absence of units is due to the current
simulation will not be tested against experimental results, where
scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}. In
\acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing
value of 0.025 $[L]$. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows that the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than \acrshort{lme}. This can be attributed to imprecision in the
stress field due to grid-crossing phenomena. Alternative to
  \acrshort{lme}, \acrshort{ugimp} shape
functions can be employed to mitigate grid-crossing instabilities.
Figure \ref{fig:Reactions-cracked-panel-mode-I} support this since it
does not exhibits it when particles cross element borders as occur
with linear interpolation. Unfortunately the extended support of this
kind of shape function interfere with nodes in both sides of the notch,
producing an undesired ``tie-effect''. This involves in a violation of
the free border boundary condition in the notch. Possible solutions to this effect
are the suppression of the elements in the notch, within a CRAMP-like
method where two pairs of nodes are employed, or reducing the mesh size. With \acrshort{lme}
approximants these artefacts are not required since it is possible to control the support
of the shape function reducing the grid-crossing and avoiding the
``tie-effect'' in the notch at the same time.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
remark clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. Although, apparently, the
differences in the obtained peaks are not extremely significant, it
can be owing to the aforementioned hookean material
employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
models will be employed, severe inaccuracies may appear in the
stress field that could affect dramatically to the final 
result. Focusing on the post failure behavior, it can be seen how
\acrshort{lme} produces soft stress field evolution even after
breaking. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: once the
energy is released dinamically by the crack, the waves propagate
along the domain.
